---
title: "Knapsack lab report"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{knapsackR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The knapsack problem
The knapsack problem is a combinatorial problem that involves creating a selection of some given objects with associated weight and value in such a way that their combined weigtht does not exceed a given limit, and their combined value is maximised. In this package we will only consider the 0-1 knapsack problem variation, where each object can either be selected *once* or *not at all*.

To get started, we will load the package in the chunk below.

```{r setup}
library(knapsackR)
```

For our examples below we will be using the following simulated objects.

```{r}
RNGversion(min(as.character(getRversion()),"3.5.3"))

##old sampler used for backward compatibility
## suppressWarnings() can be used so that the above warning is not displayed
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 2000
knapsack_objects <-
data.frame(
w=sample(1:4000, size = n, replace = TRUE),
v=runif(n = n, 0, 10000)
)
```

## The `knapsack_brute_force` function
The first method we will discuss is the **Brute force method**, which calculates the combined weights and values of all possible combinations. The combination of objects with the highest value that still respects the weight limit is guaranteed to be the optimal solution.

```{r}
# Example
#brute_force_knapsack(knapsack_objects[1:10,], 3500)

system.time(results <- brute_force_knapsack(knapsack_objects[1:10,], 3500))

results
```


## Dynamic programming

```{r}
# Example
dynamic_knapsack(knapsack_objects[1:10,], 3500)
```


## Greedy heuristic

```{r}
# Example
greedy_knapsack(knapsack_objects[1:10,], 3500)
```
